---
title: "Understanding different visualization methods using Trees, Rubber and oddbooks dataset"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Trees
Let us check the structure of trees dataset
```{r trees}
library(ggplot2)

summary(trees)
str(trees)

sapply(trees, is.factor)
```

We make use of linear model and then plot it

```{r plot }
model <- lm(Girth ~ Height, data = trees)
model1 <- lm(Volume ~ Height, data = trees)

plot(trees$Height, trees$Girth, main = 'Girth vs Height', xlab = 'Height', ylab = 'Girth', col = 'blue')

abline(model, col = 'red')

plot(trees$Height, trees$Volume, main = 'Volume vs Height', xlab = 'Height', ylab = 'Volume', col = 'blue')

abline(model1, col = 'red')

```

The same can be performed using ggplot
```{r using ggplot}
ggplot(trees, aes(Height,Girth), alpha = 0.4) + geom_point(aes(color = Volume)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

ggplot(trees, aes(Height,Volume), alpha = 0.4) + geom_point(aes(color = Girth)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

```

Histogram and density 

```{R Hist}
hist(trees$Girth, col = 'grey', xlab = "Trees", main = "Histogram of Girth", breaks = 5 ) 

hist(trees$Girth, col = 'grey', xlab = "Trees", main = "Histogram of Girth", freq = F) 
d <- density(trees$Girth)
lines(d, lwd = 2, col = 'darkblue')

curve(dnorm(x, mean=mean(trees$Girth), sd=sd(trees$Girth)), add=TRUE, col='red', lwd=2)

hist(trees$Height)
hist(trees$Volume)
```





The same can be performed using ggplot
```{R hist ggplot}
ggplot(trees, aes(Girth)) + geom_histogram(fill = 'lightblue', binwidth = 1, color = 'darkblue') + theme_bw()

ggplot(trees, aes(Girth)) + geom_histogram(aes(y=..density..) ,binwidth = 1, color="darkblue", fill="lightblue" ) + theme_bw() + geom_density(alpha=.2, fill="#FF6666")


```

Boxplot and rug

```{R b}
boxplot(trees$Girth, col = 'lightblue', horizontal = T, xlab = 'Girth', main = 'Tree Girth Data', frame.plot = F, boxwex = 0.6, notch = T)
rug(trees$Girth, side = 1)

ggplot(trees, aes(x='',y=Girth)) + geom_boxplot(color = 'darkblue', fill = 'lightblue', alpha = 0.4)+ theme_bw() + coord_flip() + ggtitle("Tree Girth Data")
```

rnorm and qqnorm

```{R rnorm}
qqnorm(trees$Girth)
qqline(trees$Girth, col = 'red')

qqnorm(trees$Volume)
qqline(trees$Volume, col = 'red')

log.Volume <- log(trees$Volume)
qqnorm(log.Volume)
qqline(log.Volume, col = 'red')

par(mfrow = c(2,4))
qqnorm(trees$Height)
qqline(trees$Height, col = 'red')
for (i in 1:7){ qqnorm(rnorm(31),xlab="", ylab="Simulated lengths",
                      main="Simulated")
qqline(rnorm(31), col = 'red')
}
```


histogram for rnorm and dnorm

```{R hist rnorm}
par(mfrow=c(1,1))
random <-  rnorm(1000,0,1)

hist(random, main="Random draw from Std Normal", cex.axis=.8, freq = F, xlim = c(-4,4))
curve(dnorm(x, mean(random), sd(random)), add=TRUE, col="darkblue", lwd=2)

```

##Rubber

Let us view and plot rubber dataset

```{R rubber}
library(MASS)

plot(Rubber)

ggplot(Rubber, aes(hard, loss)) + geom_point(aes(color = tens)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

ggplot(Rubber, aes(tens, loss)) + geom_point(aes(color = hard)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

```

We can see a negative correlation between loss and hard/tens

```{r summ}
Rubber.lm <- lm(loss~hard+tens, data=Rubber)

summary(Rubber.lm)
```

The intercept 885.16 indicates the loss at hardness = 0  and tensile strength = 0
The increase in hardness by 1 will lead to decrease in loss by 6.57
The increase in tensile strength by 1 will lead to decrease in loss by 1.37
The p value is 0.001 which indicates the confidence of 99.999

There is a negative corelation between loss and hardness/tensile strength

Let us plot the linear model using plot and termplot

```{r lm }
plot(Rubber.lm)

par(mfrow=c(1,2))
termplot(Rubber.lm, partial=TRUE, smooth=panel.smooth)
par(mfrow=c(1,1))
```


## Oddbooks
Let us install the package DAAG which consists of Oddbooks dataset

```{r od}
library(DAAG)
```

Let us plot using ggplot

```{r gg odd}
ggplot(oddbooks, aes(thick, weight)) + geom_point(aes(color = breadth)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

ggplot(oddbooks, aes(thick+ breadth, weight)) + geom_point(aes(color = breadth)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

ggplot(oddbooks, aes(thick+height+breadth, weight)) + geom_point(aes(color = thick)) + theme_bw() +geom_smooth(method = 'lm', formula = y ~x, se = F) + scale_color_continuous(high = "#FF6E2E", low = "#55D8CE")

```


Using linear model on oddbooks

```{r odd lm}
logbooks <- log(oddbooks)

logbooks.lm1 <- lm(weight~thick,data=logbooks)
summary(logbooks.lm1)$coef

```

There is a negative correlation between weight and thickness
At 0 thickness the weight is 9.69 and with the increase in Weight by 1 the thickness will decrease by 1.07 
This gives us a very weird result as generally weight should increase with increase in thickness

```{r odd lm1}
logbooks.lm2<-lm(weight~thick+height,data=logbooks)
 summary(logbooks.lm2)$coef
```
There is a negative correlation between weight and thickness/height
 At 0 thickness/height the weight is -1.26 and with the increase in Weight by 1 the thickness will increase by 0.3  and height by 2.11 
However we can see a low confidence interval for thickness and Weight 

```{r odd lm2}
logbooks.lm3<-lm(weight~thick+height+breadth,data=logbooks)
  summary(logbooks.lm3)$coef 
```

There is a negative correlation between weight and thickness/height/breadth
At 0 thickness/height the weight is -0.71 and with the increase in Weight by 1 the thickness will increase by 0.46  and height by 0.15 and breadth by 1.87 
However we can see a low confidence intervals 
  
We can see very different results from oddboks dataset as books may be 
 chosen in such a way to produce odd results
  
 
 
